\TheTitle

### 2015-08-12 linvis.tex
Architecture, Pictures, and treemap
Adding architecture section, pictures for the treemap and list viewer, and
adding more information about why the treemap is not suitable for our use-case.

Maintainers of older version of the Linux kernel must sift through thousands of
commits with tools that are unable to filter and effectively visualize projects
that are of the scale of the kernel. Tools like gitk provide users with the
full view of the DAG, showing all merges and commits. For smaller projects,
this information is useful to show the relation between various branches. With
larger, modular projects like the Linux Kernel, the DAG becomes a tangled mess
of lines that lose meaning. Github provides a DAG view for most projects, but is
unable to display the DAG of projects as large as the Linux Kernel.

@@

The second issue with performing searches is that git log, gitk, and other
tools only store the authored date of a commit. A commit may be merged at any
time, regardless of when the commit was authored. Because of this, performing
date-range queries does not provide meaningful results. Performing a date-range
query using other tools over the span of a release will contain results that
are for previous versions of the kernel, and may not contain all of the results
for that kernel if patches were merged after the release date.

@@

\subsection{Treemap visualizer}
University of Maryland presented a treemap design for displaying all the files
in a given release. This information doesn't provide meaining about what has
changed. The information within the treemap does not provide clear insight on
the meaning of the files, what changes were made, or even the size. The treemap
structure provides information on where in the directory tree the file exists.

%% https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html
%% Find paper for this maybe?

\evantodo{Picture of visualization}
% \begin{figure}[h!]
% 	\centering
% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
% 	\caption{}
% 	\label{}
% \end{figure}

\subsection{List viewer}
Moghaddam designed a graph showing how various commits in the Kernel were
linked together. This graph doesn't provide a clear picture of what is
happening though, and gives no insight to the hierarchy of the data. It
provides information about who authored the commit, and when the commit was
authored.
%% http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html
%% Find paper for this maybe?

\evantodo{Picture of graph}
% \begin{figure}[h!]
% 	\centering
% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
% 	\caption{}
% 	\label{}
% \end{figure}

@@

To build an effective tool, we focused on two main areas, the navigation
and presentation.

@@

Navigation is only effective if the destination contains the desired
information. We present the commit log, what files were edited, and what
modules were edited at a given commit or merge. Normal tools are unable to
provide information about what files are edited in a merge because they do no
use the tree structure. From our tree structure, we are able to traverse the
subtree, making note of what files are edited and how many lines are added and
removed at every commit.

>>>>

Maintainers of older version of the Linux kernel must sift through thousands of
commits with tools that are unable to filter and effectively visualize projects
that are of the scale of the kernel. Tools like gitk provide users with the
full view of the DAG, showing all merges and commits. For smaller projects,
this information is useful to show the relation between various branches.
With larger, modular projects like the Linux Kernel, the DAG becomes a tangled
mess (\ref{fig:gitk}). Github provides a DAG view for most projects, but is
unable to display the DAG of projects as large as the Linux Kernel.

@@

Another issue with performing searches is that git log, gitk, and other tools
only store the authored date of a commit. A commit may be merged at any time,
regardless of when the commit was authored. Because of this, performing
date-range queries does not provide meaningful results. Performing a date-range
query using other tools over the span of a release will contain results that
are for previous versions of the kernel, and may not contain all of the results
for that kernel if patches were merged after the release date.

@@

\subsection{Treemap visualizer}
University of Maryland presented a treemap design for displaying all the files
in a given release. Treemaps are good for displaying the topology of shallow,
wide trees, making it a good candidate for displaying file systems. Treemaps
are limited in what information can be displayed. The information within the
treemap does not provide clear insight on what changes were made to the files
in the given release. The treemap visualization only provides information on
where the file is located within the filesystem\ref{fig:treemap}. File systems
and the merge tree of the kernel share many similarities, in both cases, the
representative trees are wide and shallow, suggesting that a treemap could
potentially work to visualize the commit and merge structure of the kernel.

The treemap is unable to take into account multiple merges at the same level
with the same name. There is an additional dimension, in that the commits and
merges are ordered in time. The treemap design is unable to show the ordering
of the merges and commits.

The treemap design is also designed for working with data of differing types.
In the filesystem, there are many kinds of files.  There are text files,
scripts, images, binary files, and other types. These can be given colours
making differences stand out. There is no obvious way to assign meaningful
types to the commits such that all the commits within a given merge do not all
have the same colour.

Finally, the treemap does not provide any additional information about the
contents of each cell, only that the cell exists. The user is unable to see
their current position within the map, or if we assume that their current
position is the larges surrounding box, they can no-longer see how their
current position interacts with the rest of the system.
%% https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html
%% Find paper for this maybe?

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.47\textwidth]{figures/kernel-files.png}
	\caption{Treemap of Linux 2.5.33 file structure}
	\label{fig:treemap}
\end{figure}

\subsection{List viewer}
Moghaddam designed a web-based visualization tool displaying the chain of
merges into the kernel. There is no information on what commits are within a
given merge, or what files were edited. The visualization tool provides
information about who authored the commit and when it was authored.
\evantodo{Link to tool}
%% http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html
%% Find paper for this maybe?

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.47\textwidth]{figures/gitvis.png}
	\caption{List Viewer}
	\label{fig:listviewer}
\end{figure}

\section{Model}

We use a tree.

@@

\subsection{Architecture}

% Database
We split the data presented among 5 tables; commits, filesmod, logs,
pathtomerge, and releases. The commits table stores the metadata about a given
commit or merge. This includes the author, the authored date, the committer,
commit date, and the file-level diff. The filesmod table stores the information
about the files modified in a commit. We have the commit id, or cid, of the
commit, the filename, the number of lines added and the number of lines removed
from the file. There may be multiple files within a given commit, and there may
be multiple commits editing the same file, the database uses an index as the
primary key, though the primary key could be the cid, filename pair.

% Tables
commits
filesmod
logs
pathtomerge
releases

@@

Navigation is only effective if the destination contains the desired
information. We present the commit log, what files were edited, and what
modules were edited at a given commit or merge. Normal tools are unable to
provide information about what files are edited in a merge because they do no
use the tree structure. From our tree structure, we are able to traverse the
subtree, making note of what files are edited and how many lines of code are
added and removed at every commit.
